The Linguistic Landscape of Stream-of-Consciousness Literature

### 2024-04-03 paper/paper.qmd
update graphs

library(tidyverse)
library(tidytext)
library(rmarkdown)
library(gutenbergr)
library(tm)
library(dplyr)
library(ggplot2)
library(scales)
library(knitr)
library(wordcloud)

portrait <- read_csv(here::here("data/raw_data/portrait.csv"))
portrait_clean <- read_csv(here::here("data/analysis_data/portrait_clean.csv"))
all_books <- read_csv(here::here("data/raw_data/all_books.csv"))
combined_books <- read_csv(here::here("data/raw_data/combined_books.csv"))
portrait_words <- read_csv(here::here("data/analysis_data/portrait_words.csv"))
swann_words <- read_csv(here::here("data/analysis_data/swann_words.csv"))
dalloway_words <- read_csv(here::here("data/analysis_data/dalloway_words.csv"))
bliss_words <- read_csv(here::here("data/analysis_data/bliss_words.csv"))
prufrock_words <- read_csv(here::here("data/analysis_data/prufrock_words.csv"))
combined_trigram <- read_csv(here::here("data/analysis_data/combined_trigram.csv"))
combined_bigram <- read_csv(here::here("data/analysis_data/combined_bigram.csv"))
combined_text <- read_csv(here::here("data/analysis_data/combined_text.csv"))

@@

The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@rgutenberg]. Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `wordcloud` [@rCloud], `tm` [@rTm] and `knitr` [@rKnitr].

@@

portrait |>
  rename("Book ID" = gutenberg_id,
         "Text" = text,
         "Book" = book) |>
  tail(10) |>
  kable(
    col.names = c("Book ID", "Text", "Book"),
    booktabs = TRUE
  )

@@

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-frequency-female
#| fig-cap: "Comparative Analysis of Word Frequency in Female Stream of Consciousness Authors"

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-frequency-male
#| fig-cap: "Comparative Analysis of Word Frequency in Male Stream of Consciousness Authors"

@@

## Combined Texts: Trends, Bigram and Trigram Analsis

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-cloud
#| layout-ncol: 2
#| fig-cap: "Visualizations of Word Frequency when Combining All Stream of Consciousness Novels"
#| fig-subcap: ["One Word"] #Add "Two Words" and "Three Words" subsequently when uncovering the trends for the previous one

filtered_text <- combined_text[combined_text$n >= 300, ] #only select words with a frequency larger or equal to 300

wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors= colorRampPalette(c("lightgray", "black"))(9))

#filtered_bigram <- combined_bigram[combined_bigram$n >= 2, ]
#wordcloud(words = filtered_bigram$word, freq = filtered_bigram$n, scale=c(3,0.5), min.freq = 2, random.order = #FALSE, colors=brewer.pal(8, "Dark2"))

#filtered_trigram <- combined_trigram[combined_trigram$n >= 2, ]
#wordcloud(words = filtered_trigram$word, freq = filtered_trigram$n, scale=c(3,0.5), min.freq = 2, random.order #= FALSE, colors=brewer.pal(8, "Dark2"))

@@

# Discussion

>>>>

library(tidyverse)
library(tidytext)
library(rmarkdown)
library(gutenbergr)
library(tm)
library(dplyr)
library(ggplot2)
library(scales)
library(knitr)
library(igraph)
library(widyr)
library(ggraph)

portrait <- read_csv(here::here("data/raw_data/portrait.csv"))
portrait_clean <- read_csv(here::here("data/analysis_data/portrait_clean.csv"))
all_books <- read_csv(here::here("data/raw_data/all_books.csv"))
combined_books <- read_csv(here::here("data/raw_data/combined_books.csv"))
portrait_words <- read_csv(here::here("data/analysis_data/portrait_words.csv"))
swann_words <- read_csv(here::here("data/analysis_data/swann_words.csv"))
dalloway_words <- read_csv(here::here("data/analysis_data/dalloway_words.csv"))
bliss_words <- read_csv(here::here("data/analysis_data/bliss_words.csv"))
prufrock_words <- read_csv(here::here("data/analysis_data/prufrock_words.csv"))
combined_text <- read_csv(here::here("data/analysis_data/combined_text.csv"))

@@

The novel texts used for analysis were sourced from Project Gutenberg under the library `gutenbergr` [@rGuten] [@gutenberg]. Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `tidytext` [@rTidytext], `rmarkdown` [@rMark], `dplyr` [@rDplyr], `ggplot2` [@rGgplot2], `scales` [@rScales], `here` [@rHere], `igraph` [@rIgraph], `widyr` [@rWidy], `ggraph` [@rGg] `tm` [@rTm] and `knitr` [@rKnitr].

@@

portrait |>
  rename("Book ID" = gutenberg_id,
         "Text" = text,
         "Book" = book) |>
  head(10) |>
  kable(
    col.names = c("Book ID", "Text", "Book"),
    booktabs = TRUE
  )

@@

### Generating Word Networks

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-word-gender
#| layout-ncol: 2
#| fig-cap: "Comparative Analysis of Word Frequency in Female and Male Stream of Consciousness Authors"
#| fig-subcap: ["Female Authors", "Male Authors"]

@@

@@

## Combined Texts: Trends, Word Networks, Bigram and Trigram Analsis

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false

## Code sourced from https://www.youtube.com/watch?v=ae_XVhjHd_o

generate_word_graph <- function(combined_text,
                                minimum_n = 100,
                                minimum_correlation = 0.2) {

combined_text_count <- combined_text |>
  count(word, sort = TRUE) |>
  filter(n >= 200)

combined_correlations <- combined_text |>
  semi_join(combined_text_count, by = "word") |>
  pairwise_cor(item = word, feature = book) |>
  filter(correlation >= 0.2)

graph_from_data_frame(d = combined_correlations, #create graphs of networks of words
                      vertices = combined_text_count |>
                      semi_join(combined_correlations, by = c("word" = "item1"))) |>
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = correlation), alpha = 0.05) +
  geom_node_point() +
  geom_node_text(aes(color = n, label = name), repel = TRUE)

}

# Code sourced from
# filtered_text <- combined_text[combined_text$n >= 100, ] #only select words with a frequency larger or equal to 100

# wordcloud(words = filtered_text$word, freq = filtered_text$n, scale=c(3,0.5), min.freq = 2, random.order = FALSE, colors= colorRampPalette(c("lightgray", "black"))(9))

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-word-networks
#| layout-ncol: 2
#| fig-cap: "Word Networks Measured by Frequency and Correlation when Combining All Stream of Consciousness Novels"
#| fig-subcap: ["100 Frequency, 0.2 Correlation", "200 Frequency, 0.4 Correlation", "200 Frequency, 0.8 Correlation"] #Add "Two Words" and "Three Words" subsequently when uncovering the trends for the previous one

combined_text|>
  generate_word_graph(minimum_n = 100, minimum_correlation = 0.2)

combined_text|>
  generate_word_graph(minimum_n = 200, minimum_correlation = 0.4)

combined_text|>
  generate_word_graph(minimum_n = 200, minimum_correlation = 0.8)
```
# Discussion
